Automatic discovery of cell types and microcircuitry from neural connectomics

### 2015-04-23 docs/arxiv.tex
updates before push

@@

\printbibliography

@@

\newpage
\section*{Supplementary Methods }

\subsection*{Link Prediction}
As a proxy for link-prediction accuracy we compute the probability of a link
between two cells using each model, trained fully on the data. While this method
is potentially prone to overfitting, the overfitting will be shared across models
and in fact will preferentially bias in favor of competing models
which over-cluster the data. We use a full collection of posterior samples when
computing the link probability, and then compute the area under the ROC curve for
each.

\subsection*{Model Comparison}
We compare our model with a standard network clustering model, the
latent-position clustering model. This model assumes each cell belongs
to one of K clusters, and each cluster is associated with a
$d-$dimensional Gaussian distribution. The probability of a link is
then a function of the distance between the data points in this
continuous-space. We use \autocite{Salter-Townshend2013} a variational
implementation provided in R, parametrically varying the number of
latent dimensions and the number of requested groups.  While this
model provides reasonable link predictive accuracy, the clusterings
dramatically disagree with those from human anatomists.

\subsection* {Mouse Retina}
\label{supp:mouseretina}
Dense serial electron microscopy of a $114\mu m \times 80 \mu m $ area
in the mouse retina by \autocite{Helmstaedter2013} yielded a listing
of places where neurons come into contact. There were over 1000 cells
originally, and selected the $950$ for which the location of the soma
could be reconstructed from the provided cell plots (soma locations
were not provided by the study's authors in machine-readable
form). Ultimately this left a matrix between the total synapse-like
contact area between all pairs of 950 cells. Area was thresholded at
$0.1\mu m$, determined by hand, to yield a 950 $\times$ 950 entry
matrix that served as input to our algorithm. We measured the distance
between cells using the reconstructed soma centers, and used the
Logistic-Distance spatial relation. Hyperprior griddings are shown in
supplemental section \ref{supp:hyperpriors}.

\subsection*{C. elegans}

\subsection*{Microprocessor}
We extracted the connection graph for the transistors in the MOS6502
\autocite{visual6502source}. Each transistor has three terminals (gate,
source, drain), but the methods of the original dataset were unable to
consistently resolve which of the C1 and C2 terminals were source and
drain, leading to ambiguity in our encoding. We identified a region
consisting of three registers X, Y, and S via visual
inspection and focused our efforts there. We created a total of six
connectivity matrices by examining possible terminal pairings. One
graph, for example, $R^{gc_1}(e_i, e_j)=1$ if transistor $e_j$ and
$e_j$ are connected via pins $g$ and $c_1$.

% \begin{addendum}
%  \item Put acknowledgements here.
%  \item[Competing Interests] The authors declare that they have no
% competing financial interests.
%  \item[Correspondence] Correspondence and requests for materials
% should be addressed to A.B.C.~(email: myaddress@nowhere.edu).
% \end{addendum}
\newpage
\section*{Supplemental Material}

\section*{Measuring clustering similarity}
We compare discovered types to known types via cluster comparison
metrics: cluster homogeneity, cluster completeness, and the adjusted
Rand index. Homogeneity measures how many true types are in a given
found type. If every cell type is split into two types, each subtype
is still completely homogeneous. Completeness measures how many
members of a given true type are split across found types.

>>>>

\subsection*{Link Prediction}
To compute link-prediction accuracy we compute the probability of a
link between two cells using each model, trained via 10-fold cross
validation.  We use a full collection of posterior samples when computing
the link probability, and then compute the area under the ROC curve
for each.

We compare our model with a standard network clustering model, the
latent-position clustering model. This model assumes each cell belongs
to one of K clusters, and each cluster is associated with a
$d-$dimensional Gaussian distribution. The probability of a link is
then a function of the distance between the data points in this
continuous-space. We use \autocite{Salter-Townshend2013} a variational
implementation provided in R, parametrically varying the number of
latent dimensions and the number of requested groups.  While this
model provides reasonable link predictive accuracy, the clusterings
dramatically disagree with those from human anatomists.

\subsection*{Dataset details}
\subsubsection* {Mouse Retina}
\label{supp:mouseretina}
Dense serial electron microscopy of a $114\mu m \times 80 \mu m $ area
in the mouse retina by \autocite{Helmstaedter2013} yielded a listing
of places where neurons come into contact. There were over 1000 cells
originally, and selected the $950$ for which the location of the soma
could be reconstructed from the provided cell plots (soma locations
were not provided by the study's authors in machine-readable
form). Ultimately this left a matrix between the total synapse-like
contact area between all pairs of 950 cells. Area was thresholded at
$0.1\mu m$, determined by hand, to yield a 950 $\times$ 950 entry
matrix that served as input to our algorithm. We measured the distance
between cells using the reconstructed soma centers, and used the
Logistic-Distance spatial relation. Hyperprior griddings are shown in
supplemental section \ref{supp:hyperpriors}.

\subsubsection*{C. elegans}

\subsubsection*{Microprocessor}
We extracted the connection graph for the transistors in the MOS6502
\autocite{visual6502source}. Each transistor has three terminals (gate,
source, drain), but the methods of the original dataset were unable to
consistently resolve which of the C1 and C2 terminals were source and
drain, leading to ambiguity in our encoding. We identified a region
consisting of three registers X, Y, and S via visual
inspection and focused our efforts there. We created a total of six
connectivity matrices by examining possible terminal pairings. One
graph, for example, $R^{gc_1}(e_i, e_j)=1$ if transistor $e_j$ and
$e_j$ are connected via pins $g$ and $c_1$.

% \begin{addendum}
%  \item Put acknowledgements here.
%  \item[Competing Interests] The authors declare that they have no
% competing financial interests.
%  \item[Correspondence] Correspondence and requests for materials
% should be addressed to A.B.C.~(email: myaddress@nowhere.edu).
% \end{addendum}

\subsection*{Measuring clustering similarity}
We compare discovered types to known types via cluster comparison
metrics: cluster homogeneity, cluster completeness, and the adjusted
Rand index. Homogeneity measures how many true types are in a given
found type. If every cell type is split into two types, each subtype
is still completely homogeneous. Completeness measures how many
members of a given true type are split across found types.

@@

\printbibliography
\newpage

@@
